---
title: "my_reproducible_simulation"
author: 
  - name: Luca Y. Kogelheide
    orcid: 0009-0000-6724-0779
    email: l.y.kogelheide@uu.nl
    affiliations:
      - name: Methodology & Statistics @ UU University (Student)
date: 16 Jan 2025
date-format: "D MMM YYYY"
bibliography: Docs/references.bib
format: pdf
---

# R Code
### *Of a small simulation study that investigates the impact of measurement error in linear regression analysis.*


### Load Libraries:
```{r message=FALSE, warning=FALSE}
if(!require(Hmisc)){install.packages('Hmisc')}
library(Hmisc)
if(!require(mice)){install.packages('mice')}
library(mice)
if(!require(tidyverse)){install.packages('tidyverse')}
library(tidyverse)
```

### Read in the data:
```{r}
d1 <- sasxport.get("Data/DEMO_I.XPT")
d2 <- sasxport.get("Data/BPX_I.XPT")
d3 <- sasxport.get("Data/BMX_I.XPT")
d4 <- sasxport.get("Data/GHB_I.XPT")
d5 <- sasxport.get("Data/TCHOL_I.XPT")

###Select variables of interest
d1.t <- subset(d1, select = c("seqn","riagendr","ridageyr"))
d2.t <- subset(d2, select = c("seqn","bpxsy1"))
d3.t <- subset(d3, select = c("seqn","bmxbmi"))
d4.t <- subset(d4, select = c("seqn","lbxgh"))
d5.t <- subset(d5, select = c("seqn","lbdtcsi"))

###Merge data sets
d <- merge(d1.t,d2.t)
d <- merge(d,d3.t)
d <- merge(d,d4.t)
d <- merge(d,d5.t)
```

### Rename Variables:
- RIAGENDR: Gender
- RIDAGEYR: Age in years at screening
- BPXSY1: Systolic: Blood pres (1st rdg) mm Hg
- BMXBMI: Body Mass Index (kg/m**2)
- LBDTCSI: Total Cholesterol (mmol/L)
- LBXGH: Glycohemoglobin (%)
```{r}
d$age <- d$ridageyr
d$sex <- d$riagendr
d$bp <- d$bpxsy1
d$bmi <- d$bmxbmi
d$HbA1C <- d$lbxgh
d$chol <- d$lbdtcsi
d$age[d$age<18] <- NA
```

### Select Only Complete Cases:
```{r}
dc <- cc(subset(d,select=c("age","sex","bmi","HbA1C","bp")))
```

### Analysis:
```{r}
summary(lm(bp ~ HbA1C + age + as.factor(sex), data=dc))
confint(lm(bp ~ HbA1C + age + as.factor(sex), data=dc))
summary(lm(bp ~ HbA1C + bmi + age + as.factor(sex), data=dc))
confint(lm(bp ~ HbA1C + bmi + age + as.factor(sex), data=dc))
```


### Simulation of Measurement Error:
```{r}
ref <- lm(bp ~ HbA1C + bmi + age + as.factor(sex), data=dc)$coef[2] 
n.sim <- 1e3 
perc.me.exp <- seq(0,.5,.1)
perc.me.conf<- seq(0,.5,.1)
scenarios <- expand.grid(perc.me.exp,perc.me.conf)
var.exp <- var(dc$HbA1C)
var.conf <- var(dc$bmi)
n <- dim(dc)[1]
beta.hat <- matrix(ncol=dim(scenarios)[1], nrow=n.sim)
for (k in 1:n.sim){
  print(k)
  set.seed(k)
  for (i in 1:dim(scenarios)[1]){
    var.me.exp <- var.exp*scenarios[i,1]/(1-scenarios[i,1])
    var.me.conf <- var.conf*scenarios[i,2]/(1-scenarios[i,2])
    dc$HbA1C.me <- dc$HbA1C + rnorm(dim(dc)[1], 0, sqrt(var.me.exp) )
    dc$bmi.me <- dc$bmi + rnorm(dim(dc)[1], 0, sqrt(var.me.conf) )
    beta.hat[k,i] <- lm(bp ~ HbA1C.me + age + bmi.me + as.factor(sex),
                        data=dc)$coef[2]
  }}
```

### Create Figure with Results
```{r}
tot.mat <- cbind(100*scenarios,apply(beta.hat,2,mean))
colnames(tot.mat) <- c("me.exp","me.conf","estimate")
FIGURE <- ggplot(tot.mat, aes(me.exp, me.conf)) +
  geom_tile(color="white",aes(fill = estimate)) +
  geom_text(aes(label = round(estimate, 2))) +
  scale_fill_gradient2(low="#D55E00",mid="white",high = "#56B4E9", midpoint=ref) +
  labs(x=paste("% of total variance of HbA1c due to measurement error"),
       y=paste("% of total variance of BMI due to measurement error")) +
  coord_equal()+
  scale_y_continuous(breaks=unique(tot.mat[,1]))+
  scale_x_continuous(breaks=unique(tot.mat[,1]))+
  theme(panel.background = element_rect(fill='white', colour='grey'),
        plot.title=element_text(hjust=0),
        axis.ticks=element_blank(),
        axis.title=element_text(size=12),
        axis.text=element_text(size=10),
        legend.title=element_text(size=12),
        legend.text=element_text(size=10))
FIGURE
```

### Save Figure: 
#### *Please note, this figure is already included in the folder called "Results". Therefore, this step is only needed when changes to the code were made.
```{r}
ggsave(filename = "Results/Figure_STRATOS.tif", plot = FIGURE, device = "tiff")
```

## End of R Code